---
title: "Week 03 - Map Classification"
author: "Adela Sobotkova"
date: "06/01/2021 updated`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

##  Learn about Classification methods

This exercise demonstrates some possibilities of classifications in R, based on Ch.9 of [Geocomputation](https://geocompr.robinlovelace.net/adv-map.html)

### Required R libraries

* We will use the sf, raster, and tmap packages.
* Additionally, we will use the spData and spDataLarge packages that provide new datasets.

```{r}
library(sf)
library(raster)
library(tmap)
library(spData)
library(spDataLarge)
```


### Data sets 

We will use a single data set:  `nz`. It is contained by the libraries
It is an sf object with polygons representing the 16 regions of New Zealand.

### Existing code

Here are some examples of plotting population in  New Zealand. * Your role is to create a map based on the suggestions below, selecting the most meaningful classification style.

```{r nz-pop}
# Look at NZ population distribution
hist(nz$Population)
```


```{r maps}
# This line of code applies the 'pretty' style rounding legend numbers. Try different numbers of classes.
a <- tm_shape(nz) + tm_polygons(col = "Population", style = "pretty", n = 4)

# "Jenks" style further smooths over the gaps
b <- tm_shape(nz) + tm_polygons(col = "Population", style = "jenks", n = 5)

# quantile style divides into 5 even groups
c <- tm_shape(nz) + tm_polygons(col = "Population", style = "quantile", n=5)

# Equal interval style divides the distribution into even groups
d <- tm_shape(nz) + tm_polygons(col = "Population", style = "equal", n = 5)

# Write maps above to objects and plot them side by side 
# with tmap_arrange() for better comparison
tmap_arrange(a,b,c,d)

```



#### Questions

* 1. What are the advantages and disadvantages of each classification method?
Some of the classification methods group the populations two broadly - giving less detail. If all areas are marked as 

* 2. Choose the best classification and create a map with easily legible legend and all other essentials.(Select a suitable color palette from http://colorbrewer2.org/, north arrow, scale, map title, legend title, reasonable number of breaks in the classification )

* 3. Which method and how many classes did you end up using to display your data? Why did you select that method?

We ended up selecting the jenks interval scheme as it had more detail than the first plot !

NOTES: (From https://mgimond.github.io/Spatial/symbolizing-features.html): 
"The Jenks interval scheme (aka natural breaks) uses an algorithm that identifies clusters in the dataset. The number of clusters is defined by the desired number of intervals."

* 4. What principles did you use in arranging the parts of your map layout the way you did and how did you apply these principles?
Using sequential color palette to show increasing size in population. Placing the 

# Your solution

```{r}
# /Start Code/ #
#define color palette
palette = c('#fee5d9','#fcbba1','#fc9272','#fb6a4a','#de2d26','#a50f15')

# make map 
b <- tm_shape(nz) + 
    tm_polygons(col = "Population", style = "jenks", n = 5, palette = palette) +
    
    tm_compass(position = c("left", "top"),
             type = "rose", 
             size = 3) +
    
    tm_layout(
            main.title = "Population of New Zealand",
            main.title.position = "center",
            inner.margins = c(0.1, 0.1, 0.1, 0.1), 
            legend.position=c("right", "bottom"),
            bg.color = "lightblue"
            ) 

b

pdf("nzmap.pdf")
b 
#tmap_arrange(a,b,c,d)
dev.off()

# /End Code/ #
```

